---
title: "Finding proxies for energy consumption by exploratory data analysis (EDA), and test if they correlate with installtion of renewable energy"
author: "Yangguang Li"
date: "10/07/2017"
output:
  html_document:
    fig_caption: yes
  html_notebook: default
---
``` {r load_libraries, message=FALSE, warning = FALSE}
setwd("..")
# Loading the environment
library(ProjectTemplate)
load.project()
```
```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.path="../graphs/")
```
The analyses were done on two data sets: Living Cost and Food Survey (LCF) and Understanding Society (UK Household Longitudinal Study, UKHLS).

## Living Cost and Food Survey (LCF)

### Data set
We combined two data sets of households and grouped individual expenditures for each household in dv_set to be the sample used. 

However, the variables we chosed to scrutinise from data set dv_set, 5.3.1.3.2 (elec_cookers), 5.3.1.4.1 (heaters), 5.5.1.1.1 (elec_tools), 5.5.2.1.4 (elec_consumables), 9.1.1.1.1 (audio_equipment), and 9.1.1.1.3 (accessories_audio_equip), were not available for all households, and after tranforming and filtering out the observations that had missing values, we had `r 0` left. Also all households had reported energy consumption did not use renewable energy, making the `r nrow(lcf_explo)` observations available had no positive responses.

Hence it was impossible to do the analysis on LCF.

The process of munging data can be found at [munge_data LCF](../munge/01-lcf-data.R).

## Understanding Society (UK Household Longitudinal Study, UKHLS)

### Data set
We used the wave d of UKHLS (deposited Nov. 2014) for the analysis. All the variables below have been removed the prefix 'd_'. The sample had `r nrow(d_hh)` observations, and among them `r nrow(d_hh) - nrow(ukhls_explo)` were discarded (`r nrow(ukhls_explo)` left) as they had missing values.

First we found the proxies for energy consumption through EDA, and then tested if they were correlated with the installtion of renewable energy.

Dependent variables:  
1. xpduely (amount electricity and gas), xpelecy (amount electricity), xpgasy (amount gas), xpoily (amount oil), and xpsfly (amount other fuel), are sumed as **energy** to estimate the energy consumption of the household.  
2. solar1 (install solar panels for electricity), solar2 (install solar water heating), and solar3 (install wind turbine to generate electricity) are grouped into single variable **renewable** which is TRUE if any of them was answered yes.

There were three classes of independent variables were scrutinized, including:  
1. solar1, solar2, and solar3  
2. cduse1 through to cduse13 (possesion of 13 kinds of appliances, detailed in the table beneath)  
3. fuelduel, duelpay, xpelecy, elecpay, xpgasy, gaspay (consumption data about energy, detailed in the table beneath)  

#### Munging
``` {r ukhls_explo_variables, echo = FALSE}
tribble(
  ~Former, ~Tranformed, ~Categories,
  "xpduely, xpelecy, xpgasy, xpoily, and xpsfly", "energy (£/yr)", writeContinuous(ukhls_explo$energy),
  "solar1, solar2, solar3", "renewable", writeCategories(ukhls_explo %>% count(renewable)),
  "solar1", "solar_elec", writeCategories(ukhls_explo %>% count(solar_elec)),
  "solar2", "solar_heating", writeCategories(ukhls_explo %>% count(solar_heating)),
  "solar3", "wind_tur", writeCategories(ukhls_explo %>% count(wind_tur)),
  "cduse1", "color_tv", writeCategories(ukhls_explo %>% count(color_tv)),
  "cduse2", "video_recorder_or_dvd_player", writeCategories(ukhls_explo %>% count(video_recorder_or_dvd_player)),
  "cduse3", "satellite_dish_or_sky_tv", writeCategories(ukhls_explo %>% count(satellite_dish_or_sky_tv)),
  "cduse4", "cable_tv", writeCategories(ukhls_explo %>% count(cable_tv)),
  "cduse5", "fridge_freezer", writeCategories(ukhls_explo %>% count(fridge_freezer)),
  "cduse6", "washing_machine", writeCategories(ukhls_explo %>% count(washing_machine)),
  "cduse7", "tumble_dryer", writeCategories(ukhls_explo %>% count(tumble_dryer)),
  "cduse8", "dish_washer", writeCategories(ukhls_explo %>% count(dish_washer)),
  "cduse9", "microwave", writeCategories(ukhls_explo %>% count(microwave)),
  "cduse10", "pc", writeCategories(ukhls_explo %>% count(pc)),
  "cduse11", "disc_player", writeCategories(ukhls_explo %>% count(disc_player)),
  "cduse12", "telephone", writeCategories(ukhls_explo %>% count(telephone)),
  "cduse13", "mobile", writeCategories(ukhls_explo %>% count(mobile)),
  "fuelduel", "separate_or_combined_bills", writeCategories(ukhls_explo %>% count(separate_or_combined_bills)),
  "duelpay", "payment_method", writeCategories(ukhls_explo %>% count(payment_method)),
  "xpelecy", "elec_amount (£/yr)", writeContinuous(ukhls_explo$elec_amount),
  "elecpay", "elec_pay_method", writeCategories(ukhls_explo %>% count(elec_pay_method)),
  "xpgasy", "gas_amount (£/yr)", writeContinuous(ukhls_explo$gas_amount),
  "gaspay", "gas_pay_method", writeCategories(ukhls_explo %>% count(gas_pay_method))
) %>% kable()
```

The process of munging data can be found at [munge_data UKHLS](../munge/05-ukhls-data.R).  
``` {r UKHLS_explo_data, echo = FALSE}
ukhls_explo %>% head(5) %>% select(hidp:satellite_dish_or_sky_tv) %>% kable()
ukhls_explo %>% head(5) %>% select(hidp,cable_tv:telephone) %>% kable()
ukhls_explo %>% head(5) %>% select(hidp,mobile:gas_pay_method) %>% kable()
```

### Analysis

The analysis script can be found at [analysis_proxies UKHLS](../munge/08-ukhls-proxies.R)

#### EDA: finding proxies
We first fitted for each independent variable a linear model predicting the energy consumption, then depending on the adjusted $R^2$ of those models and plots of energy consumption versus the corresponding variables, we found the following proxies: tumble_dryer (adjusted $R^2$ = `r writePercent(glance(ukhls_proxy_tumble_dryer)[[2]])`%), dish_washer (`r writePercent(glance(ukhls_proxy_dish_washer)[[2]])`%), pc (`r writePercent(glance(ukhls_proxy_pc)[[2]])`%), elec_amount (`r writePercent(glance(ukhls_proxy_elec_amount)[[2]])`%), and gas_amount (`r writePercent(glance(ukhls_proxy_gas_amount)[[2]])`%). All of the models' p values were <= 0.01.
``` {r proxy_adj_r.sq_and_plot, echo = FALSE}
plot1 <- ukhls_explo %>% ggplot(aes(energy, ..density..)) + geom_freqpoly(aes(color = tumble_dryer), binwidth = 300)
plot2 <- ukhls_explo %>% ggplot(aes(energy, ..density..)) + geom_freqpoly(aes(color = dish_washer), binwidth = 300)
plot3 <- ukhls_explo %>% ggplot(aes(energy, ..density..)) + geom_freqpoly(aes(color = pc), binwidth = 300)
grid.arrange(plot1, plot2, plot3, nrow = 3)
ukhls_explo %>% ggplot(aes(gas_amount, energy)) + geom_point(position = "jitter")
ukhls_explo %>% ggplot(aes(elec_amount, energy)) + geom_point(position = "jitter")
```

Then we fitted the model with energy consumption as dependent variables, and independent variabls being the proxies found.

`energy ~ tumble_dryer + dish_washer + pc + elec_amount + gas_amount`   
The variables combined`r hasSignificance(ukhls_proxy_all)` explained `r writePercent(glance(ukhls_proxy_all)[[1]])`% of the variability in energy consumption (adjusted $R^2$ = `r writePercent(glance(ukhls_proxy_all)[[2]])`%), F(`r glance(ukhls_proxy_all)[["df"]]`, `r glance(ukhls_proxy_all)[["df.residual"]]`) = `r writeNum(glance(ukhls_proxy_all)[["statistic"]])`, p = `r writePValue(glance(ukhls_proxy_all)[["p.value"]])`. None of the VIFs were greater than 5. Detailed regression result can be found at [Table A](#table-a).
``` {r EDA_all, echo = FALSE}
vif(ukhls_proxy_all) %>% as.tibble() %>% rename(VIF = value) %>% kable()
```

And then we tested the correlation between all proxies and installation of renewable energy.

#### Test correlation
Since the dependent variables were all binary, we used logistic regression for fitting the models, and we used McFadden's pseudo $R^2 = 1 - \frac{ln(\hat{L}_M)}{ln(\hat{L}_0)}$ where $\hat{L}_M$ and $\hat{L}_0$ are likelihoods for the model being fitted and the null model for analysis.

1. `renewable ~ all proxies`  
The model fitted`r hasSignificance(ukhls_proxy_result)` had pseudo $R^2$ of `r writeNum(analysis_Logit(ukhls_proxy_result)[[1]])` and p = `r writePValue(analysis_Logit(ukhls_proxy_result)[[2]])`. All VIFs were smaller than 5. Detailed regression result can be found at [Table B](#table-b).
``` {r ukhls_proxy_result, echo = FALSE}
vif(ukhls_proxy_result) %>% as.tibble() %>% rename(VIF = value) %>% kable()
```

2. `solar_elec ~ all proxies`  
The model fitted`r hasSignificance(ukhls_proxy_solar_elec_result)` had pseudo $R^2$ of `r writeNum(analysis_Logit(ukhls_proxy_solar_elec_result)[[1]])` and p = `r writePValue(analysis_Logit(ukhls_proxy_solar_elec_result)[[2]])`. All VIFs were smaller than 5.

3. `solar_heating ~ all proxies`  
The model fitted`r hasSignificance(ukhls_proxy_solar_heating_result)` had pseudo $R^2$ of `r writeNum(analysis_Logit(ukhls_proxy_solar_heating_result)[[1]])` and p = `r writePValue(analysis_Logit(ukhls_proxy_solar_heating_result)[[2]])`. All VIFs were smaller than 5.

4. `wind_tur ~ all proxies`  
The model fitted`r hasSignificance(ukhls_proxy_solar_heating_result)` had pseudo $R^2$ of `r writeNum(analysis_Logit(ukhls_proxy_solar_heating_result)[[1]])` and p = `r writePValue(analysis_Logit(ukhls_proxy_solar_heating_result)[[2]])`. All VIFs were smaller than 5.

## Appendix

##### Table A   
Results of the OLS regression. Here the table, B means unstandardised coefficient, β standardised coefficient. All proxies.
``` {r table-a, echo = FALSE}
temp <- tribble(
  ~Predictor,
  "Tumble Dryer",
  "Dish Washer",
  "PC",
  "Amount Electricity",
  "Amount Gas",
  "Intercept"
) 
temp %>% cbind(tidy.beta(ukhls_proxy_all)) %>%
  rename(B = Estimate, β = Standardized, SE = `Std. Error`) %>% 
  mutate(Predictor = predictor_add_stars(Predictor, `Pr(>|t|)`)) %>%
  select(Predictor, B, SE, β) %>%
  kable()
```

Signif.codes: \*\*\* 0.001, \*\* 0.01, * 0.05

##### Table B
Results of the logistic regression. All proxies.
``` {r table-b, echo = FALSE}
temp %>% cbind(tidy(ukhls_proxy_result)) %>%
  rename(B = estimate, SE = std.error, z = statistic, `P(>|z|)` = p.value) %>% 
  mutate(Predictor = predictor_add_stars(Predictor, `P(>|z|)`)) %>%
  select(-term) %>%
  kable()
```
